---
title: "Linegraph Lab"
author: "Randy Treat"
date: "2023-02-15"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# This is the setup chunk
#  Here you can set global options for the entire document

library(knitr)
library(ggplot2) # I recommend doing this

# Although you can call functions from a library using the following notation
#  without loading the entire library.
knitr::opts_chunk$set(echo = TRUE, 
                      comment = NA, # Recommended
                      fig.path = "./figures/",  # Store all figures here in relative path (make the folder first)
                      fig.align = "center",
                      fig.width = 7, # default figure width in inches
                      fig.height = 7, # default figure height in inches
                      message = FALSE, # Turn off load messages
                      warning = FALSE # Turn off warnings
                      )

```


```{r include=FALSE}
# You should not echo this chunk.
# include=FALSE does more than echo=FALSE, it actually does: echo=FALSE, results='hide', 
#  message=FALSE, warning=FALSE, and fig.show='hide'

# You should set your working directory at the very beginning of your R Markdown file

# You should type things out properly for your system
# Easiest Way - Session>Set Working Directory>To Source Files Location
# THEN, Copy/paste from R Console here 

```

<!-- ---------------------------------------------------------------------------------------------------- -->
<!-- ---------------- Lab starts below these lines ----------------------------------------- -->
<!-- ---------------------------------------------------------------------------------------------------- -->



***

## Line Graphs

Using the housing dataset, we want to see the trends of the house values over time.

Filter only the first Quarter from each year (`Qrtr == 1`) when considering the different years and discard all other rows. Use either the "filter" function (install the dplyr library) or "subset" function from Base R to accomplish this. 

Do all of the following:

i. Have different colored lines for different states.

ii. Make sure the x-axis has ticks starting with 1975 and ending with 2013 with ticks every 5 years in-between.

iii. Make sure the y-axis has ticks for every 100,000 (starting with 100,000 and ending with 800,000) but format it in a dollar amount $XXX,XXX. Use the dollar_format() function from the scales package to accomplish this.  

iv. Change the y-axis label to one that does not have a period between the words.

v. Add an appropriate title.

```{r, echo=TRUE}
setwd("C:/Users/tomse/OneDrive/Desktop/Data Viz")
library(dplyr)
library(scales)
housing<-read.csv("housing.csv", header=T)
housing_qrtr <- housing%>%
                  filter(Qrtr==1)
 all_states <- ggplot(housing_qrtr) +
  aes(x = Year, y = Home.Value, colour = State) +
  geom_line() +
  scale_color_hue(direction = 1) +
  theme_minimal()+
  scale_y_continuous("Home Value", breaks = seq(0, 800000, by=100000),labels=scales::dollar_format())+
  scale_x_continuous("Year", breaks = seq(1975, 2013, by= 5))+
   ggtitle("Rising and Falling Home Values")
 
 all_states
  
```

Now, color all lines gray except for California. Make California red. Add a nicer theme.

```{r}
CA_red <- all_states + geom_line(data = housing, aes(group = State), color = "grey") +
    scale_color_manual(values = c("CA" = "red"))

CA_red
```

What happened around 2008 that might be relevant to the dip in home prices everywhere? Was California more or less affected during this dip in prices?

_Your Answer_
Mortgage market crash, recession
